---
title: "Chapter 5 & 6"
author: "Tong Sun"
date: "2/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.width="0.9\\linewidth",dev="pdf",fig.align  = 'center')
```
## 5.8
We will now perform cross-validation on a simulated data set.
###(a)
Generate a simulated data set as follows. In this data set, what is n and what is p? Write out the model used to generate the data in equation form.
```{r echo=FALSE}
set.seed(1)
y <- rnorm(100)
x <- rnorm(100)
y <- x - 2 * x^2 + rnorm(100)
```
Here we have that n = 100 and p = 2, the model used is $Y = X - 2X^2 + \epsilon$.

###(b)
Create a scatterplot of X against Y. Comment on what you find.
```{r echo=FALSE}
plot(x, y)
```
I find that the data obviously holds a curved relationship.

###(c)
Set a random seed, and then compute the LOOCV errors that result from fitting the following four models using least squares:
i. $Y=\beta_0+\beta_1X+\epsilon$
```{r echo=FALSE}
library(boot)
set.seed(1)
model1 <- data.frame(x, y)
glm.1 <- glm(y ~ x)
cv.glm(model1 ,glm.1) $ delta[1]
```


ii. $Y=\beta_0+\beta_1X+\beta_2X^2+\epsilon$
```{r echo=FALSE}
# 'poly' function returns or evaluates orthogonal polynomials of degree 1 to degree over the specified set of points 'x'
glm.2 <- glm(y ~ poly(x, 2))
cv.glm(model1, glm.2)$delta[1]
```

iii. $Y=\beta_0+\beta_1X+\beta_2X^2+\beta_3X^3+\epsilon$
```{r echo=FALSE}
glm.3 <- glm(y ~ poly(x, 3))
cv.glm(model1, glm.3)$delta[1]
```

iv. $Y=\beta_0+\beta_1X+\beta_2X^2+\beta_3X^3+\beta_4X^4+\epsilon$
```{r echo=FALSE}
glm.4 <- glm(y ~ poly(x, 4))
cv.glm(model1, glm.4)$delta[1]
```

Note you may find it helpful to use the data.frame() function to create a single data set containing both X and Y.

###(d)
Repeat (c) using another random seed, and report your results. Are your results the same as what you got in (c)? Why?
```{r echo=FALSE}
set.seed(100)
glm.1 <- glm(y ~ x)
cv.glm(model1 ,glm.1) $ delta[1]
```

```{r echo=FALSE}
glm.2 <- glm(y ~ poly(x, 2))
cv.glm(model1, glm.2)$delta[1]
```

```{r echo=FALSE}
glm.3 <- glm(y ~ poly(x, 3))
cv.glm(model1, glm.3)$delta[1]
```

```{r echo=FALSE}
glm.4 <- glm(y ~ poly(x, 4))
cv.glm(model1, glm.4)$delta[1]
```

I find that both of these two seeds above have the same results because LOOCV evaluates n folds of a single observation.
###(e)
Which of the models in (c) had the smallest LOOCV error? Is this what you expected? Explain your answer.

I find that the LOOCV estimates from (c) for the lowest test MSE is "glm.2", which has quadratic predictor in the equation. I also see that in (b) that the relationship between "x" and "y" is quadratic. So I think this result of model fitting is considerable.

###(f)
Comment on the statistical significance of the coefficient estimates that results from fitting each of the models in (c) using least squares. Do these results agree with the conclusions drawn based on the cross-validation results?

```{r echo=FALSE}
summary(glm.4)
```

The p-values show that the linear and quadratic terms are statistically significant , which have the p-values lower than 0.05, and that the cubic and 4th degree terms are not statistically significant. Therefore, this result agrees with the conclusions drawn based on the cross-validation results.

## 6.2
For parts (a) through (c), indicate which of i through iv is correct. Justify your answer.
###(a)
The lasso, relative to least squares, is:
i. More flexible and hence will give improved prediction accuracy when its increase in bias is less than its decrease in variance.
ii. More flexible and hence will give improved prediction accuracy when its increase in variance is less than its decrease in bias.
iii. Less flexible and hence will give improved prediction accuracy when its increase in bias is less than its decrease in variance.
iv. Less flexible and hence will give improved prediction accuracy when its increase in variance is less than its decrease in bias.

The third one is right. The lasso, relative to least squares, is less flexible and hence will give improved prediction accuracy when its increase in bias is less than its decrease in variance. Lasso's advantage over least squares is the bias-variance trade-off. When the least squares estimates have excessively high variance, the lasso solution can yield a reduction in variance at the expense of a small increase in bias. Therefore, it can generate more accurate predictions. The other advantage of lasso is that it performs variable selection which makes it easier to interpret than other methods like ridge regression.

###(b)
Repeat (a) for ridge regression relative to least squares.

Also the third one is right. The ridge regression relative to least squares, is less flexible and hence will give improved prediction accuracy when its increase in bias is less than its decrease in variance.The same as Lasso, Ridge regression's advantage over least squares is the bias-variance trade-off. As $\lambda$ increases, the flexibility of the ridge regression fit decreases leading to decreased variance but increased bias. Considering the relationship between $\lambda$ and variance and bias in different regression methods: when there is small change in the training data, the least squares coefficient produces a big change and a larger value of variance as well. But for ridge regression, it can still perform well by trading off a small increase in bias for a large decrease in variance so that the test MSE will not get larger too much. Therefore, between these two methods, ridge regression works better in  the situation where the least squares estimates have high variance.

###(c)
Repeat (a) for non-linear methods relative to least squares.

The second one is right. The non-linear methods relative to least squares, is more flexible and hence will give improved prediction accuracy when its increase in variance is less than its decrease in bias.

## 6.9
In this exercise, we will predict the number of applications received using the other variables in the College data set.
###(a)
Split the data set into a training set and a test set.
```{r echo=FALSE}
library(ISLR)
data(College)
set.seed(1)
data1 <- sample(1:dim(College)[1], dim(College)[1] / 2)
data2 <- -data1
train <- College[data1, ]
test <- College[data2, ]
```
Here I got a training set named "train" and a test set named "test".

###(b)
Fit a linear model using least squares on the training set, and report the test error obtained.
```{r echo=FALSE}
fit1 <- lm(Apps ~ ., data = train)
pred.lm <- predict(fit1, test)
mean((pred.lm - test$Apps)^2)
```
Using least squares on the training set, the test error obtained is 1135758.

###(c)
Fit a ridge regression model on the training set, with $\lambda$ chosen by cross-validation. Report the test error obtained.
```{r echo=FALSE}
library(glmnet)
train.mat <- model.matrix(Apps ~., data = train[ ,-1])
test.mat <- model.matrix(Apps ~., data = test[ ,-1])
set.seed(1)
cv.out <- cv.glmnet(train.mat, train$Apps, alpha = 0)
bestlam <- cv.out$lambda.min
bestlam
```

```{r echo=FALSE}
pred.ridge <- predict(cv.out, s = bestlam, newx = test.mat)
mean((pred.ridge - test$Apps) ^2)
```
It would be better to use cross-validation to choose the tuning parameter $\lambda$. Here I set a random seed first so the results will be reproducible, since the choice of the cross-validation folds is random. Therefore, I see that the value of $\lambda$ that results in the smallest cross-validation error is 405.8404. This time I want to get predictions for a test set, by using "newx" argument. And the test MSE is 1007688, which represents a further improvement over the test MSE that I got using least squares.

###(d)
Fit a lasso model on the training set, with $\lambda$ chosen by cross-validation. Report the test error obtained, along with the number of non-zero coefficient estimates.
```{r echo=FALSE}
set.seed(1)
cv.out <- cv.glmnet(train.mat, train$Apps, alpha = 1)
plot(cv.out)
bestlam.lasso <- cv.out$lambda.min
bestlam.lasso
pred.lasso <- predict(cv.out, s = bestlam.lasso, newx = test.mat)
mean((pred.lasso - test$Apps) ^2)
```
We can see from the coefficient plot that depending on the choice of tuning parameter, some of the coefficients will be exactly equal to zero. Next I perform cross-validation and compute the associated test error, which is 1140473 here. So I think the Lasso model is not as strong at predicting as the ridge regression on this data.

```{r echo=FALSE}
lasso.coef<- predict(cv.out, s = bestlam.lasso, type = "coefficients")[1:18, ]
lasso.coef
lasso.coef[lasso.coef != 0]
```
However, the lasso has a substantial advantage over ridge regression in that the resulting coefficient estimates are sparse. Here I find that 7 of the 18 coefficient estimates are exactly zero. So the lasso model with $\lambda$ chosen by cross-validation contains only 11 variables.

###(e)
Fit a PCR model on the training set, with M chosen by cross-validation. Report the test error obtained, along with the value of M selected by cross-validation.
```{r echo=FALSE}
library(pls)
pcr.fit <- pcr(Apps ~ ., data = train, scale = TRUE, validation = "CV")
summary(pcr.fit)
validationplot(pcr.fit, val.type = "MSEP")
```

Here I apply PCR to the data.Setting "scale = TRUE" has the effect of standardizing each predictor, prior to generating the principal components, so that the scale on which each variable is measured will not have an effect. Setting 'validation ="CV"' causes "pcr()" to compute the ten-fold cross-validation error for each possible value of M, the number of principal components used.The CV score is provided for each possible number of components, ranging from M=0 onwards. One can also plot the cross-validation scores using the "validationplot" function. We see that the smallest cross-validation error occurs when  about M =17 components are used.

```{r echo=FALSE}
pred.pcr <- predict(pcr.fit, test.mat, ncomp = 5)
mean((pred.pcr - test$Apps)^2)
```

This test set MSE is 1963819 here.

###(f)
Fit a PLS model on the training set, with M chosen by cross-validation. Report the test error obtained, along with the value of M selected by cross-validation.
```{r echo=FALSE}
set.seed(1)
pls.fit <- plsr(Apps ~ ., data = train, scale = TRUE, validation = "CV")
summary(pls.fit)
validationplot(pls.fit, val.type = "MSEP")
pred.pls <- predict(pls.fit, test.mat, ncomp = 10)
mean((pred.pls - test$Apps) ^2)
```
 The test MSE here in partial least squares is 1181808.

###(g)
Comment on the results obtained. How accurately can we predict the number of college applications received? Is there much difference among the test errors resulting from these five approaches?

In order to compare the results obtained above, we need to compute the test $R^2$ for all models.
```{r echo=FALSE}
test.avg <- mean(test$Apps)
lm <- 1- mean((pred.lm - test$Apps)^2) / mean((test.avg - test$Apps)^2)
ridge <- 1- mean((pred.ridge - test$Apps)^2) / mean((test.avg - test$Apps)^2)
lasso <- 1- mean((pred.lasso - test$Apps)^2) / mean((test.avg - test$Apps)^2)
pcr <- 1- mean((pred.pcr - test$Apps)^2) / mean((test.avg - test$Apps)^2)
pls <- 1- mean((pred.pls - test$Apps)^2) / mean((test.avg - test$Apps)^2)
lm
ridge
lasso
pcr
pls
```

So the test $R^2$ for the least squares is 0.9015413, the test $R^2$ for ridge regression is 0.9015558, the test $R^2$ for lasso regression is 0.9011326, the test $R^2$ for PCR is 0.8297569, and the test $R^2$ for PLS is 0.8975493. Maybe I can conclude that, except PCR, all models predict college applications with high accuracy.

## 6.10
We have seen that as the number of features used in a model increases, the training error will necessarily decrease, but the test error may not. We will now explore this in a simulated data set.
###(a)
Generate a data set with p = 20 features, n = 1,000 observations, and an associated quantitative response vector generated according to the model $Y=X\beta +\epsilon$, where $\beta$ has some elements that are exactly equal to zero.
```{r echo=FALSE}
set.seed(1)
x <- matrix(rnorm(1000 * 20), 1000, 20)
beta <- rnorm(20)
beta[1] <- 0
beta[5] <- 0
beta[10] <- 0
beta[15] <- 0
beta[20] <- 0
epsil <- rnorm(1000)
y <- x %*% beta + epsil
```

###(b)
Split your data set into a training set containing 100 observations and a test set containing 900 observations.
```{r echo=FALSE}
part1 <- sample(seq(1000), 100, replace = FALSE)
part2 <- -part1
x.train <- x[part1, ]
x.test <- x[part2, ]
y.train <- y[part1]
y.test <- y[part2]
```

###(c)
Perform best subset selection on the training set, and plot the training set MSE associated with the best model of each size.
```{r echo=FALSE}
data1 <- data.frame(y = y.train, x = x.train)
library(leaps)
regfit.full <- regsubsets(y ~ ., data = data1, nvmax = 20)
reg.summary <- summary(regfit.full)
reg.summary$rsq
train.mat <- model.matrix(y ~ ., data = data1, nvmax = 20)
error.train <- rep(NA, 20)
for(i in 1 :20){
  coef <- coef(regfit.full, id = i)
  pred <- train.mat[, names(coef)] %*% coef
  error.train[i] <- mean((pred - y.train) ^ 2)
}
plot(error.train, xlab = "Number of Predictors", ylab = "Training MSE", pch = 16, type = "b")
```
"regsubsets()" function performs best subset selection by identifying the best model that contains a given number of predictors, where best is quantified using RSS.  An asterisk indicates that a given variable is included in the corresponding model.  For instance, the output above shows that the best two-variable model contains only "x.9" and "x.4". The "nvmax" option can be used in order to return as many variables as are desired. Here I fit uo to a 20-variable model. The "summary()" function also returns $R^2$, RSS, adjusted $R^2$, $C_p$, and BIC. For example, I find that the $R^2$ statistics increases from 33.6%, when only one variable is included in the model, to almost 95.2%, when all variables are included. As expected, the $R^2$ statistic increases monotonically as more variables are included. Next I make a model matrix from the training data. And then I run a loop, and for each size $i$, I extract the coefficients from "regfit.full" for the best model of that size, multiply them into the appropriate columns of the training model matrix to form the predictions, and compute the training MSE. I find that the best model is the one that contains all of the variables.

###(d)
Plot the test set MSE associated with the best model of each size.
```{r echo=FALSE}
data2 <- data.frame(y = y.test, x = x.test)
regfit.best <- regsubsets(y ~ ., data = data2, nvmax = 20)
best.summary <- summary(regfit.best)
best.summary$rsq
test.mat <- model.matrix(y ~ ., data = data2, nvmax = 20)
error.test <- rep(NA, 20)
for(i in 1 :20){
  coef <- coef(regfit.full, id = i)
  pred <- test.mat[, names(coef)] %*% coef
  error.test[i] <- mean((pred - y.test) ^ 2)
}
plot(error.test, xlab = "Number of Predictors", ylab = "Test MSE", pch = 16, type = "b")
```
Here I do the same thing on test data. The $R^2$ statistics increases from 22.7%, when only one variable is included in the model, to almost 94.0%, when all variables are included. And then after viewing the test MSE, I find that the best model is the one that contains 15 of the variables.

###(e)
For which model size does the test set MSE take on its minimum value? Comment on your results. If it takes on its minimum value for a model containing only an intercept or a model containing all of the features, then play around with the way that you are generating the data in (a) until you come up with a scenario in which the test set MSE is minimized for an intermediate model size.
```{r echo=FALSE}
which.min(error.test)
```
The 15-variables model has the smallest test MSE, which can be concluded by the test MSE in (d).

###(f)
How does the model at which the test set MSE is minimized compare to the true model used to generate the data? Comment on the coefficient values.
```{r echo=FALSE}
coef(regfit.full, which.min(error))
```
The best model caught all zeroed out coefficients.

###(g)
Create a plot displaying $\sqrt{\sum_{j=1}^{p}(\beta_j-\hat \beta^{r}_{j})^2}$ for a range of values of $r$  where $\hat \beta^{r}_{j}$ is the jth. coefficient estimate for the best model containing $r$ coefficients. Comment on what you observe. How does this compare to the test MSE plot from (d)?
```{r echo=FALSE}
error <- rep(NA, 20)
x_cols <- colnames(x, do.NULL = FALSE, prefix = "x.")
for (i in 1:20) {
  coef <- coef(regfit.full, id = i)
  error[i] <- sqrt(sum((beta[x_cols %in% names(coef)] - coef[names(coef) %in% x_cols])^2) + sum(beta[!(x_cols %in% names(coef))])^2)
}
plot(error, xlab = "Number of coefficients", ylab = "Error between estimated and true coefficients", pch = 19, type = "b")
```
We may see that the model with 3 variables minimizes the error between the estimated and true coefficients. However test error is minimized by the model with 15 variables. So, a better fit of true coefficients does not necessarily mean a lower test MSE.

## 6.11
We will now try to predict per capita crime rate in the Boston data set.
###(a)
Try out some of the regression methods explored in this chapter, such as best subset selection, the lasso, ridge regression, and PCR. Present and discuss results for the approaches that you consider.

# Best subset selection
```{r echo=FALSE}
library(MASS)
data(Boston)
set.seed(1)
library(leaps)
predict.regsubsets <- function(object, newdata, id, ...){
  form <- as.formula(object$call[[2]])
  mat <- model.matrix(form, newdata)
  coefi <- coef(object, id = id)
  mat[, names(coefi)] %*% coefi
}

k = 10
p <- ncol(Boston) - 1
folds <- sample(rep(1:k, length = nrow(Boston)))
cv.errors <- matrix(NA, k, p)
for(i in 1:k){
  best.fit <- regsubsets(crim ~ ., data = Boston[folds != i, ], nvmax = p)
  for(j in 1:p) {
    pred <- predict(best.fit, Boston[folds == i, ], id = j)
    cv.errors[i,j] <- mean((Boston$crim[folds == i] - pred)^2)
  }
}
rmse.cv <- sqrt(apply(cv.errors, 2, mean))
plot(rmse.cv, pch = 16, type = "b")
```

Here I try to choose among the models of different sizes using cross-validation. First, we create a vector that allocates each observation to one of k=10 folds, and I create a matrix in which I will store the results. Then I write a for loop that performs cross-validation. In the $i$th fold, the elements of folds that equal $i$ are un the test set, and the remainder are in the training set. I make the predictions for each model size, compute the test errors on the appropriate subset, and store them in the appropriate slot in the matrix "cv.errors". The code will automatically use the "predict.regsubsets()" function when I call "predict()" because the "best.fit" object has class "regsubsets". After fitting 13 models, which equals to the number of variables minus 1, I will need to find the one model that minimizes the CV error on the test data.
```{r echo=FALSE}
summary(best.fit)
```

```{r echo=FALSE}
which.min(rmse.cv)
```

```{r echo=FALSE}
boston.bsm.err <- (rmse.cv[which.min(rmse.cv)]) ^2
boston.bsm.err
```

I see that cross-validation selects a 9-variable model based on the test MSE. At 9-variables, the CV estimate for the test MSE is 43.32807. The variables that are included in this model are zn,indus,nox,dis,rad,ptratio,black,lstat, and medv.

# The lasso
```{r echo=FALSE}
x <- model.matrix(crim ~ ., Boston)[, -1]
y <- Boston$crim
boston.lasso <- cv.glmnet(x, y, alpha = 1, type.measure = "mse")
plot(boston.lasso)
```
The graph above depicts the relationship between log $\lambda$ and MSE. To help predict the training model on the test model, I will need to find the $\lambda$ that reduces the error the most.
```{r echo=FALSE}
coef(boston.lasso)
```
```{r echo=FALSE}
boston.lasso.err <- (boston.lasso$cvm[boston.lasso$lambda == boston.lasso$lambda.1se])
boston.lasso.err
```

As we know that Lasso is a variable reduction method. From the results shown above, the Lasso model that reduces the MSE the model includes only one variable and has an MSE of 55.02399. The only variable included in this model is "rad".
# Ridge regression
```{r echo=FALSE}
boston.ridge <- cv.glmnet(x, y, alpha = 0, type.measure = "mse")
plot(boston.ridge)
```
Ridge regression keeps all the variables but push their coefficient value close to zero if they do not have significance in the relationship with the response.
```{r echo=FALSE}
coef(boston.ridge)
```
```{r echo=FALSE}
boston.ridge.err <- boston.ridge$cvm[boston.ridge$lambda == boston.ridge$lambda.1se]
boston.ridge.err
```
The MSE for the ridge regression method is 61.37358 -- much larger than those in other two methods. So I think the ridge regression doesn't perform well.
# PCR
```{r echo=FALSE}
library(pls)
boston.pcr <- pcr(crim ~ ., data = Boston, scale = TRUE, validation = "CV")
summary(boston.pcr)
```
Based on the CV error as well as the variances explained, I think that the appropriate PCR model would only include 8 components. With 8 components, 93.45% of the variance is explained in the predictors by the model, and 42.47% of the variance is explained in the response variable by the model. Additionally, at 8 components, the MSE is at

###(b)
Propose a model (or set of models) that seem to perform well on this data set, and justify your answer. Make sure that you are evaluating model performance using validation set error, cross-validation, or some other reasonable alternative, as opposed to using training error.

As mentioned above, the model that has the lowest cross-validation error is the one chosen by the best subset selection method, which has a MSE of 43.32807.
###(c)
Does your chosen model involve all of the features in the data set? Why or why not?

The model that was chosen by Best Subset Selection only includes 9 variables. The variables that are included in this model are zn, indus, nox, dis, rad, ptratio, black, lstat and medv. If the model were to include of the thrown-out features, more variation of the response would be present. For this particular problem, we are looking to have model prediction accuracy with low variance and low MSE.